---
title: "Reproducible Research Assignment - 1"
output: html_document
---

## Loading and preprocessing the data

### Load the data

```{r}
# The downloaded file activity.csv is placed in working directory
getwd()
x <- read.csv("activity.csv")
dim(x)
```

## What is mean total number of steps taken per day?

### Make a histogram of the total number of steps taken each day

```{r}
# Aggregate total steps by date
aggdata <-aggregate(x$steps,by=list(x$date),FUN=sum,na.rm=TRUE)

# Create Histogram 
hist(aggdata$x,main="Histogram of Total Number of Steps",xlab="Steps",col = "RED")
```

### Calculate and report the mean and median total number of steps taken per day

```{r}

# Mean of total number of steps
aggregate(x$steps,by=list(x$date),FUN=mean,na.rm=TRUE)

# Median of total number of steps
aggregate(x$steps,by=list(x$date),FUN=median,na.rm=TRUE)
```

## What is the average daily activity pattern?

### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
# Aggregate data for average steps by interval
aggdata <-aggregate(x$steps,by=list(x$interval),FUN=mean,na.rm=TRUE)

# Generate time series plot
plot(aggdata$Group.1,aggdata$x,type="l",xlab="Interval",ylab="Average Steps Taken")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
# Aggregate data for average steps by interval
aggdata <-aggregate(x$steps,by=list(x$interval),FUN=mean,na.rm=TRUE)

#Calulate maximum steps interval 
aggdata[which.max(aggdata$x),]
# The interval is 835
```

## Imputing missing values

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(x))
# The total number of missing values is 2304
```

### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
# Get the mean value for each day
aggdata <-aggregate(x$steps,by=list(x$date),FUN=mean,na.rm=TRUE)
```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
names(aggdata) <- c("date", "Mean")
# Merge x(original data set) and aggdata(data set with mean values for each day) to add a column Mean
xxx <- merge(x,aggdata,by="date")

# Replace all NA values of "steps" from calculated "Mean" column where "Mean" is NOT "NA"
xxx$steps[is.na(xxx$steps)] <- xxx$Mean[is.na(xxx$Mean)]

# Replace all NA values of "steps" from calculated "Mean" column where "Mean" is "NA" with overall mean values for those dates where all values are NA like "2012-10-01"
xxx[is.na(xxx)] <- mean(xxx$steps,na.rm=TRUE)

# Make equivalent data set with all NA values replaced
dataset <- data.frame(xxx$steps,xxx$date,xxx$interval)
names(dataset) <- c("steps", "date","interval")
dim(dataset)
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
# Aggregate data for average steps by date for new data set
aggdata <-aggregate(dataset$steps,by=list(dataset$date),FUN=sum,na.rm=TRUE)
hist(aggdata$x,main="Histogram of Total Number of Steps",xlab="Steps",col = "BLUE")

# New Mean of total number of steps 
aggregate(dataset$steps,by=list(dataset$date),FUN=mean,na.rm=TRUE)

# New Median of total number of steps
aggregate(dataset$steps,by=list(dataset$date),FUN=median,na.rm=TRUE)

```

## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
# Add new column "day" which has two values “weekday” and “weekend” based on day of week
dataset$day <- ifelse(weekdays(as.Date(dataset$date)) %in% c("Saturday","Sunday"),"weekend", "weekday")
```

### Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
# Aggregate data for average steps by day and interval for new data set
aggdata <-aggregate(dataset$steps,by=list(dataset$interval,dataset$day),FUN=mean,na.rm=TRUE)
names(aggdata) <- c("interval", "day","steps")                    

library(ggplot2)
ggplot(aggdata, aes(x = interval, y = steps, group = day)) +
  geom_line() +
  facet_wrap( ~ day,nrow=2) 

# Create HTML and md file using knitr package from R console
# knit2html("PA1_template.Rmd")

```



